---
title: "Report - Bayesian Subset Selection"
author: "Ezequiel"
date: "2024-10-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load libraries
library(hdbayes)
library(matrixStats)
library(tidyverse)
library(lme4)
library(loo)
library(knitr)
library(kableExtra)

# load data
data <- read_csv("data/chapman.csv")

# set parameters
ncores <- max(1, parallel::detectCores() - 2)

# load formulas
formulas <- readRDS("data/formulas.rds")
covariates_models <- readRDS("data/covariates_models.rds")
```

## Data

We used the Chapman Data, which contains the following variables:

- Ag: age of in years;
- S: systolic blood pressure in millimeters of mercury;
- D: diastolic blood pressure in millimeters of mercury;
- Ch: cholesterol in milligrams per DL;
- H: height in inches;
- W: weight in pounds;
CNT: 1 if a coronary incident had occurred in the previous ten years, and 0 otherwise.

```{r}
head(data)
```

Firstly, we created a list of formulas to be used in the Bayesian subset selection:

```{r}
head(formulas)
```

For each formula, we fit a Bayesian logistic regression model using the `glm.post` function from the `hdbayes` package. We used 2000 iterations for warmup and 2000 iterations for sampling.

```{r}
# run hdbayes
fits_bayes <- lapply(formulas, function(x)
  glm.post(x,
           family = binomial(link = "logit"),
           data = list(data),
           parallel_chains = ncores,
           iter_warmup = 2000,
           iter_sampling = 2000,
           refresh = 0,
           ))
```

And we also fit a frequentist logistic regression model using the `glm` function from the `stats` package.

```{r}
# run glm
fits_freq <- lapply(formulas, function(x) 
  glm(x, family = binomial(link = "logit"), data = data))
```

## Model Comparison

We get the log marginal likelihood with the `glm.logml.post` function from the `hdbayes` package and compute the AIC and BIC for each frequentist model. 

```{r, results=FALSE}
# compute marginal likelihood
marginal_likelihoods <- lapply(fits_bayes, function(x) glm.logml.post(x)$logml)
# compute AIC
AICs <- lapply(fits_freq, function(x) x$aic)
# compute BIC
BICs <- lapply(fits_freq, function(x) BIC(x))

# create data frame
metrics <- data.frame(
  variables = unlist(covariates_models),
  logml = unlist(marginal_likelihoods),
  AIC = unlist(AICs),
  BIC = unlist(BICs)
)

# sort by logml
metrics <- metrics[order(metrics$logml, decreasing = TRUE), ]
```

```{r}
head(metrics)
```

## Weights

We compute the weights based on the log marginal likelihood, AIC, and BIC. For the log marginal likelihood, we use the `logSumExp` function from the `matrixStats` package to avoid numerical issues and compute the weights as $w_i = \exp(\text{logml}_i - \text{logSumExp}(\text{logml}))$. For the AIC and BIC, we compute the weights as $w_i = \exp(-\text{AIC}_i - \text{logSumExp}(-\text{AIC}))$ and $w_i = \exp(-\text{BIC}_i - \text{logSumExp}(-\text{BIC}))$, respectively.

```{r}
# compute weights
weights_logml <- exp(unlist(metrics$logml) - logSumExp(unlist(metrics$logml)))
weights_AIC <- exp(unlist(-metrics$AIC) - logSumExp(unlist(-metrics$AIC)))
weights_BIC <- exp(unlist(-metrics$BIC) - logSumExp(unlist(-metrics$BIC)))

# create a data frame with the weights
weights <- data.frame(
  variables = metrics$variables,
  logml = weights_logml,
  AIC = weights_AIC,
  BIC = weights_BIC
)
```

And we get the following weights ordered by log marginal likelihood:

```{r echo=FALSE, results='asis'}
kable(weights,
      format = "latex",
      caption = "Weights based on log marginal likelihood, AIC, and BIC.",
      row.names = TRUE,
      longtable = TRUE) %>%
      kable_styling(latex_options = "HOLD_position")
```

Then, best models based on log marginal likelihood are:

```{r echo=FALSE, results='asis'}
kable(weights[1:10,],
      format = "latex",
      caption = "Best weights based on log marginal likelihood.",
      row.names = TRUE,
      longtable = TRUE) %>%
      kable_styling(latex_options = "HOLD_position")
```

## Plots

We plot the weights based on the log marginal likelihood, AIC, and BIC.

```{r echo=FALSE}
# plot
ggplot(weights, aes(x = reorder(variables, -logml), y = logml)) +
  geom_bar(stat = "identity") +
  labs(title = "Weights based on log marginal likelihood",
       x = "Variables",
       y = "Weights") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r echo=FALSE}
# plot
ggplot(weights, aes(x = reorder(variables, -AIC), y = AIC)) +
  geom_bar(stat = "identity") +
  labs(title = "Weights based on AIC",
       x = "Variables",
       y = "Weights") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r echo=FALSE}
# plot
ggplot(weights, aes(x = reorder(variables, -BIC), y = BIC)) +
  geom_bar(stat = "identity") +
  labs(title = "Weights based on BIC",
       x = "Variables",
       y = "Weights") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r echo=FALSE}
# plot: logml vs AIC
ggplot(weights, aes(x = logml, y = AIC)) +
  geom_point() +
  labs(title = "Log marginal likelihood vs AIC - Weights",
       x = "Log marginal likelihood",
       y = "AIC")
```

```{r echo=FALSE}
# plot
ggplot(weights, aes(x = logml, y = BIC)) +
  geom_point() +
  labs(title = "Log marginal likelihood vs BIC - Weights",
       x = "Log marginal likelihood",
       y = "BIC")
```

