---
title: "Estimating the ATE"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here we describe how to use the model probabilities to estimate the average treatment effect (ATE) as in section 3.3.2 of the paper.
First, we define auxiliary functions to compute marginal means for treatment arms and apply them across models via Bayesian model averaging (BMA).

```{r aux_ate_functions, cache = TRUE}
# function to compute predictions for a given treatment arm
predict_ <- function(data, outcome, trt, family, beta_post, arm){
  if (arm == 1) {
    data[[trt]] <- 1
  } else if (arm == 0) {
    data[[trt]] <- 0
  }
  cov_data <- colnames(data)[!(colnames(data) %in% c(outcome))]
  prod <- lapply(beta_post,
                 function(post_samp) {
                   cov <- names(post_samp)[names(post_samp) %in% cov_data]
                   post_samp <- as.data.frame(post_samp[c("(Intercept)", cov)])
                   cov_mat <- cbind(1, data.matrix(data[cov]))
                   cov_mat <- t(cov_mat)
                   beta <- data.matrix(post_samp)
                   beta %*% cov_mat
                 }
  )
  pred <- lapply(prod, function(df) {
    apply(df, 2, family$linkinv)
  })
  return(pred)
}

# function to compute mean outcome for a given treatment arm
mean_models_arm <- function(data, outcome, trt, family, beta_post, arm){
  pred <- predict_(data, outcome, trt, family, beta_post, arm)
  mean_models <- lapply(pred,
                        function(x) {
                          apply(x, 1, function(row){
                            sample_weights <- t(rdirichlet(1, rep(1, nrow(data))))
                            row %*% sample_weights
                          })
                        }
  )
  return(do.call(cbind, mean_models))
}

# function to compute BMA samples
bma <- function(x, df_post, n_samples = 1000){
  ks <- sample(1:nrow(df_post), n_samples, replace = TRUE, prob = df_post$post_model)
  samples <- lapply(ks, function(k) {
    id_draw <- sample(1:nrow(x), 1) 
    draw_x <- x[id_draw,k]
  })
  return(do.call(rbind, samples))
}
```

Then we define the setup for the analysis.

```{r setup_analysis, cache = TRUE, warning=FALSE}
# load libraries
library(hdbayes)
library(parallel)
library(matrixStats)
library(MCMCpack)
library(bayestestR)
library(ggplot2)
library(dplyr)
library(patchwork)
library(reshape2)

# define data
load("../samples/post_samples.RData")
current_data <- actg036

# normalize data
current_data$age <- (current_data$age - mean(current_data$age)) /
  (2*sd(current_data$age))
current_data$cd4 <- (current_data$cd4 - mean(current_data$cd4)) /
  (2*sd(current_data$cd4))

# set parameters
family <- binomial(link = "logit")
```

Now we are ready to run the procedure to compute posterior samples for the ATE. 

```{r run_ate_analysis, cache = TRUE, warning=FALSE}
mean_models_ctrl <- mean_models_arm(current_data,
                                    "outcome",
                                    "treatment",
                                    family,
                                    post_samples$post_betam,
                                    0)
mean_models_trt <- mean_models_arm(current_data,
                                   "outcome",
                                   "treatment",
                                   family,
                                   post_samples$post_betam,
                                   1)
# compute BMA
bma_ctrl <- bma(mean_models_ctrl, post_samples$df_post, 10000)
bma_trt <- bma(mean_models_trt, post_samples$df_post, 10000)
```

Finally, we can plot the results.

```{r plot_ate_results, cache = TRUE, warning=FALSE, echo=FALSE, fig.width=8.4, fig.height=4.8}
# Define colors
blended_rgb <- round(colMeans(rbind(
  c(135, 206, 235),
  c(70, 130, 180)
)))
blended_color <- rgb(blended_rgb[1], blended_rgb[2], blended_rgb[3], maxColorValue = 255)
stats_colors <- c("90% \nBCI" = blended_color,
                  "Density" = "skyblue",
                  "ATE = 0" = "black",
                  "OR = 1" = "black"
)

# Create a data frame for the ATE
df_bma <- data.frame(
  value = bma_trt - bma_ctrl
)
# Compute credibility intervals
ci_bma_95 <- bayestestR::ci(df_bma$value, ci = 0.95)
ci_bma_90 <- bayestestR::ci(df_bma$value, ci = 0.90)
ci_ate_95_lower <- ci_bma_95$CI_low
ci_ate_95_upper <- ci_bma_95$CI_high
ci_ate_90_lower <- ci_bma_90$CI_low
ci_ate_90_upper <- ci_bma_90$CI_high

# Compute density for the entire dataset
ate_density_data <- density(df_bma$value)
ate_density_df <- data.frame(x = ate_density_data$x, y = ate_density_data$y) 

# Filter density data for the HDI regions
ate_density_95 <- ate_density_df %>% filter(x >= ci_ate_95_lower & x <= ci_ate_95_upper)
ate_density_90 <- ate_density_df %>% filter(x >= ci_ate_90_lower & x <= ci_ate_90_upper)

# Compute the mean, median, and quartiles
mean_value <- mean(df_bma$value)
median_value <- median(df_bma$value)
q1_value <- quantile(df_bma$value, 0.25)
q3_value <- quantile(df_bma$value, 0.75)

# Plot using ggplot2
bma_ate <- ggplot() +
  geom_area(data = ate_density_df %>% filter(x <= ci_ate_90_lower + 0.0003), 
            aes(x = x, y = y, fill = "Density"), color = "black",
            alpha = 0.7) +
  geom_area(data = ate_density_df %>% filter(x >= ci_ate_90_upper - 0.0003), 
            aes(x = x, y = y, fill = "Density"), color = "black",
            alpha = 0.7) +
  geom_area(data = ate_density_90, 
            aes(x = x, y = y, fill = "90% \nBCI"), color = "black",
            alpha = 0.7) +
  geom_vline(aes(xintercept = 0, color = "ATE = 0"), linetype = "solid", size = 1) + 
  scale_fill_manual(name = NULL, values = stats_colors, breaks = c("90% \nBCI")) +
  scale_color_manual(name = NULL, values = stats_colors, guide = NULL) +
  labs(title = "",
       x = "Average treatment effect (ATE)", y = "", color = "") +
  theme_bw() +
  theme(legend.position = c(.95, .95),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.margin = margin(6, 6, 6, 6),
        legend.background = element_rect(fill = "white", color = "black"),
        panel.background = element_rect(fill = "white", color = NA),
        plot.background = element_rect(fill = "white", color = NA)
  ) +
  theme(text = element_text(size = 16),        
        axis.title = element_text(size = 18),  
        axis.text = element_text(size = 16),   
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 16))

# create BMA data frames
df_bma_arm <- data.frame(
  value = c(bma_ctrl, bma_trt),
  group = c(rep("ctrl", length(bma_ctrl)), rep("trt", length(bma_trt)))
)
# create OR data frame
df_or <- data.frame(or = bma_trt / (1-bma_trt) / bma_ctrl / (1-bma_ctrl))

# compute credibility intervals for the odds ratio
ci_or_90 <- bayestestR::ci(df_or$or, ci = 0.90)
ci_or_90_lower <- ci_or_90$CI_low
ci_or_90_upper <- ci_or_90$CI_high

# Compute density for the entire dataset
or_density_data <- density(df_or$or)
or_density_df <- data.frame(x = or_density_data$x, y = or_density_data$y)

# Filter density data for the HDI regions
or_density_90 <- or_density_df %>% filter(x >= ci_or_90_lower & x <= ci_or_90_upper)

# Plot the density of the odds ratio
plot_or <- ggplot() +
  geom_area(data = or_density_df %>% filter(x <= ci_or_90_lower + 0.01), 
            aes(x = x, y = y, fill = "Density"), color = "black",
            alpha = 0.7) +
  geom_area(data = or_density_df %>% filter(x >= ci_or_90_upper - 0.01), 
            aes(x = x, y = y, fill = "Density"), color = "black",
            alpha = 0.7) +
  geom_area(data = or_density_90, 
            aes(x = x, y = y, fill = "90% \nBCI"), color = "black",
            alpha = 0.7) +
  geom_vline(aes(xintercept = 1, colour = "OR = 1"), linetype = "dotted", size = 1)  +
  labs(title = "",
       x = "Odds ratio", y = "") +
  scale_fill_manual(name = NULL, values = stats_colors, breaks = c("90% \nBCI")) +
  scale_color_manual(name = NULL, values = stats_colors, guide = NULL) +
  theme_bw() +
  theme(legend.position = c(.95, .95),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.margin = margin(6, 6, 6, 6),
        legend.background = element_rect(fill = "white", color = "black"),
        panel.background = element_rect(fill = "white", color = NA),
        plot.background = element_rect(fill = "white", color = NA)
  ) +
  theme(text = element_text(size = 16),        
        axis.title = element_text(size = 18),  
        axis.text = element_text(size = 16),   
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 16))

# put or and bma_ate together
bma_ate_or <- (bma_ate + theme(legend.position = "none")) + 
  plot_or + plot_layout(ncol = 2) &
  theme(
    axis.title = element_text(size = 14),
    legend.title = element_text(size = 12),
    legend.text  = element_text(size = 14)
  )
bma_ate_or
```
