---
title: "Prior predictive check (PPC)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ppc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_knit$set(root.dir = "~/repositories/hdbayes-tutorial")
```

Here we describe how to perform a prior predictive checks (PPC) as in the section 3.3.3 of the paper.
We start by defining useful functions and Stan models to sample from the NPP.

```{r aux_functions, cache = TRUE, warning=FALSE, message=FALSE}
source("logistic_regression/code/aux_scripts/get_stan_data.r")
source("logistic_regression/code/aux_scripts/glm_npp_lognc_wip.r")
source("logistic_regression/code/aux_scripts/glm_npp_wip.r")
source("logistic_regression/code/aux_scripts/glm_logml_npp_wip.r")
source("logistic_regression/code/aux_scripts/data_checks.r")
source("logistic_regression/code/aux_scripts/expfam_loglik.r")
```

Then we define the setup to run the PPC for the ACTG data.

```{r ppc_setup, cache = TRUE}
library(hdbayes)
library(parallel)
library(matrixStats)
library(MCMCpack)
library(bayestestR)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(grid)

current_data <- actg036
hist_data <- actg019

current_data$age <- (current_data$age - mean(current_data$age)) /
  (2*sd(current_data$age))
current_data$cd4 <- (current_data$cd4 - mean(current_data$cd4)) /
  (2*sd(current_data$cd4))
hist_data$age <- (hist_data$age - mean(hist_data$age)) /
  (2*sd(hist_data$age))
hist_data$cd4 <- (hist_data$cd4 - mean(hist_data$cd4)) /
  (2*sd(hist_data$cd4))

data <- list(current_data, hist_data)
family <- binomial(link = "logit")
iter_warmup <- 1000
iter_sampling <- 2500
formula <- outcome ~ age + treatment + cd4
a0      <- seq(0, 1, length.out = 21)
logncfun <- function(a0, ...){
  glm.npp.lognc(
      formula = formula, family = family, a0 = a0, histdata = hist_data,
      ...
    )
}
ncores <- detectCores() - 1
```

Now we run the PPC for different hyperparameter values of the Beta prior on a0.

```{r sample_prior, cache = TRUE, warning=FALSE, message=FALSE}
a0_hyper <- list(c(1,1), c(2,2), c(1,10), c(10,1))
c0 <- 1
a0.lognc <- parallel::mclapply(
  X = a0, FUN = logncfun, iter_warmup = iter_warmup,
  iter_sampling = iter_sampling, chains = 4,  beta.sd = 1,
  mc.cores = ncores
)
a0.lognc <- data.frame( do.call(rbind, a0.lognc) )

draws.a0 <- lapply(seq_along(a0_hyper), function(i) {
  delta0_val <- a0_hyper[[i]][1]
  lambda0_val <- a0_hyper[[i]][2]
  standat.norm <- get.stan.data.npp.prior.wip(
    formula        = formula,
    family         = family,
    data.list      = list(hist_data),
    a0.lognc       = a0.lognc$a0,
    lognc          =  matrix(a0.lognc$a0, ncol = 1),
    offset.list    = NULL,
    beta.sd        = c0,
    disp.mean      = NULL,
    disp.sd        = NULL,
    a0.shape1      = delta0_val,
    a0.shape2      = lambda0_val,
    a0.lower       = 0,
    a0.upper       = 1,
    wip            = 0
  )

  fit.norm <- glm_npp_prior_wip$sample(
    data = standat.norm,
    iter_warmup = iter_sampling, iter_sampling = iter_warmup, chains = 4
  )

  pars <- c("beta[1]", "beta[2]", "beta[3]", "beta[4]")
  draws.norm <- fit.norm$draws(format = 'draws_df', variables = pars) %>%
    select(all_of(pars))
})
```

And then we compute prior predictive samples for each observation in the current data.

```{r prior_predictive, cache = TRUE}
nsim <- 1000
Xnew        <- stats::model.matrix(formula, current_data)
d.sub.a0 <- lapply(seq_along(a0_hyper), function(j){
  draws.a0[[j]][sample(x = seq_len(nrow(draws.a0[[j]])), size = nsim, replace = F), ]
})
pnew.a0 <- lapply(seq_along(a0_hyper), function(j) {
  sapply(seq_len(nsim), function(i){
    beta.sim <- as.numeric(d.sub.a0[[j]][i, ])
    p <- binomial('logit')$linkinv(Xnew %*% beta.sim)
    return(p)
  })
})
```

Finally, we plot the prior predictive distributions for a given observation.

```{r ppc_plot, cache = TRUE, fig.width=8, fig.height=6, echo=FALSE}
i <- 1  # index of the observation to plot
plot <- ggplot() +
    geom_density(data = data.frame(y = pnew.a0[[1]][i, ]),
                 aes(x = y, color = "a01"),
                 linewidth = 1) +
    geom_density(data = data.frame(y = pnew.a0[[2]][i, ]),
                 aes(x = y, color = "a02"),
                 linewidth = 1) +
    geom_density(data = data.frame(y = pnew.a0[[3]][i, ]),
                 aes(x = y, color = "a03"),
                 linewidth = 1) +
    geom_density(data = data.frame(y = pnew.a0[[4]][i, ]),
                 aes(x = y, color = "a04"),
                 linewidth = 1) +
    scale_color_manual(
      name = NULL,
      values = c(
        "a01" = "black",
        "a02" = brewer.pal(9, "Set1")[7],
        "a03" = brewer.pal(9, "Set1")[8],
        "a04" = brewer.pal(9, "Set1")[9]
      ),
      labels = c(
        "a01" = bquote("Beta"*"("*.(a0_hyper[[1]][1])*", "*.(a0_hyper[[1]][2])*")"),
        "a02" = bquote("Beta"*"("*.(a0_hyper[[2]][1])*", "*.(a0_hyper[[2]][2])*")"),
        "a03" = bquote("Beta"*"("*.(a0_hyper[[3]][1])*", "*.(a0_hyper[[3]][2])*")"),
        "a04" = bquote("Beta"*"("*.(a0_hyper[[4]][1])*", "*.(a0_hyper[[4]][2])*")")
      ),
      breaks = c("a01",
                 "a02",
                 "a03",
                 "a04"
      )
    ) +
    annotation_custom(
      grid::textGrob(
        label = paste0("Obs outcome: ", current_data$outcome[i]),
        x = unit(0.81, "npc"),
        y = unit(0.96, "npc"),
        gp = grid::gpar(
          col = "black",
          fontsize = 10,
          fill = "white"
        )
      )
    ) +
    theme_bw() +
    theme(legend.position = c(0.81, 0.73),
          legend.background = element_rect(fill = "white", color = "black")) +
    theme(panel.background = element_rect(fill = "white", color = NA),
          plot.background = element_rect(fill = "white", color = NA)
    ) +
    xlim(c(0,1)) +
    theme(text = element_text(size = 10),
          axis.title = element_text(size = 10),
          axis.text = element_text(size = 10),
          legend.title = element_text(size = 10),
          legend.text = element_text(size = 10),
          strip.text = element_text(size = 10)) +
    labs(
      x = "Probability of event",
      y = "",
    )
plot
```
