---
title: "Variable Selection"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{var_sel}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here we describe the analysis perfomed in the section 3.3 of the paper.
We start by defining useful functions to compute all subsets of models:

```{r aux_functions, cache = TRUE}
# function to compute power set
powerset <- function(x) {
  sets <- lapply(1:(length(x)), function(i) combn(x, i, simplify = F))
  unlist(sets, recursive = F)
}

# function to create formula
create_formula <- function(x, outcome) {
  as.formula(paste(paste(outcome, "~", sep = " "), paste(x, collapse = "+")))
}

# function to get covariates from models
get_covariates <- function(x) {
  paste(x, collapse = ", ")
}
```

Then to compute useful quantities for all models, we define the following functions. The first one computes the prior probability of a model; the second one computes the log-normalizing constant for $a_0$; and the third one computes posterior samples of beta for a given model.

```{r model_priors, cache = TRUE}
logp0 <- function(formula,
                  ...) {
  hdbayes::glm.npp.lognc(
    formula = formula,
    ...
  )
}

# function to compute log-normalizing constant for a0
logncfun <- function(a0,
                     ...){
  hdbayes::glm.npp.lognc(
    a0 = a0,
    ...
  )
}

# function to compute posterior samples of beta for a given model
post_beta <- function(formula, 
                      c0, 
                      data,
                      family,
                      delta0,
                      lambda0,
                      iter_warmup,
                      iter_sampling) {
  a0.lognc <- lapply(a0_seq, 
                     logncfun, 
                     formula = formula, 
                     family = family, 
                     histdata = data[[2]], 
                     iter_warmup = iter_warmup,
                     iter_sampling = iter_sampling, 
                     chains = 4, 
                     refresh = 0,
                     beta.sd = c0
  )
  a0.lognc <- data.frame(do.call(rbind, a0.lognc))
  fit <- hdbayes::glm.npp(formula, 
                          data = data,
                          family = family,
                          a0.lognc = a0.lognc$a0,
                          lognc = matrix(a0.lognc$lognc, ncol = 1),
                          a0.shape1 = delta0,
                          a0.shape2 = lambda0,
                          iter_warmup = iter_warmup,
                          iter_sampling = iter_sampling, 
                          chains = 4, 
                          refresh = 0,
                          beta.sd = c0
  )
}

# function to compute posterior model probabilities
post_models <- function(logp0_models, logml_models) {
  logp0_models <- sapply(logp0_models, function(x) x[2])
  logml_models <- unlist(data.frame(do.call(rbind, logml_models))$logml)
  post_model <- exp(logml_models + logp0_models -
                      logSumExp(logml_models + logp0_models))
}
```

Finally, we put everything together to compute the posterior probabilities of all models. The *logp0_models* object contains the log-marginial likelihoods of all models based on historical data $D_0$, i.e., $Z_m(D_0) = \int L(\boldsymbol{\beta}^{(m)} | D_0) \pi_0 \left(\boldsymbol{\beta}^{(m)} \mid d_0 \right) d\beta^{(m)}$; the *post_betam* object contains the posterior samples of beta for all models; the *logml_models* object contains the log-marginal likelihoods of all models, i.e., $Z_m(D) \equiv p(D | m) = \int L(\boldsymbol{\beta}^{(m)} | D) \pi\left(\beta^{(m)}, a_0 \mid D_0\right) d\beta^{(m)} da_0$; and the *df_post_ord* object contains a data frame with prior probabilities, marginal likelihoods, and posterior probabilities of all models ordered by posterior probability.

```{r post_models, cache = TRUE}
samples_models <- function(data,
                        outcome,
                        family,
                        a0_seq,
                        c0,
                        d0,
                        delta0,
                        lambda0,
                        iter_warmup,
                        iter_sampling,
                        num_cores = 10) {
  current_data <- data[[1]]
  hist_data <- data[[2]]

  covariates <- colnames(current_data)[colnames(current_data) != outcome]
  pset <- powerset(covariates)
  formulas <- lapply(pset, create_formula, outcome = outcome)
  
  covariates_models <- lapply(pset, get_covariates)
  
  cl <- parallel::makeCluster(num_cores)
  on.exit(parallel::stopCluster(cl))
  parallel::clusterExport(cl, varlist = c('a0_seq',
                                          'family',
                                          'data',
                                          'c0',
                                          'd0',
                                          'delta0',
                                          'lambda0',
                                          'iter_warmup',
                                          'iter_sampling',
                                          'logncfun'),
                          envir = environment()
  )
  
  logp0_models <- parLapply(
    cl = cl,
    X = formulas,
    fun = logp0,
    family = family,
    histdata = hist_data,
    a0 = 1,
    iter_warmup = iter_warmup,
    iter_sampling = iter_sampling,
    chains = 4,
    refresh = 0,
    beta.sd = d0
  )

  post_betam  <- parLapply(
    cl = cl,
    X = formulas,
    fun = post_beta,
    c0 = c0, 
    data = data,
    family = family,
    delta0 = delta0,
    lambda0 = lambda0,
    iter_warmup = iter_warmup,
    iter_sampling = iter_sampling
  )

  logml_models <- parLapply(
    cl = cl,
    X = post_betam,
    fun = hdbayes::glm.logml.npp
  )

  prior_models <- exp(sapply(logp0_models, function(x) x[2]) -
                        logSumExp(sapply(logp0_models, function(x) x[2]))
  )

  post_models <- post_models(logp0_models, logml_models)

  df_post <- data.frame(model = unlist(covariates_models),
                        prior_model = prior_models,
                        ml = exp(unlist(data.frame(do.call(rbind, logml_models))$logml)),
                        post_model = post_models
  )

  df_post_ord <- df_post[order(df_post$post_model, decreasing = TRUE),]
  rownames(df_post_ord) <- 1:nrow(df_post_ord)
  return(list(logp0_models = logp0_models,
              post_betam = post_betam,
              logml_models = logml_models,
              df_post = df_post,
              df_post_ord = df_post_ord))
}
```

Now we can run the analysis for the example in the paper. We start by loading the libraries and the data.

```{r load_data, cache = TRUE}
# load libraries
library(hdbayes)
library(parallel)
library(matrixStats)
library(MCMCpack)

# define data
current_data <- actg036
hist_data <- actg019

# normalize data
current_data$age <- (current_data$age - mean(current_data$age)) /
  (2*sd(current_data$age))
current_data$cd4 <- (current_data$cd4 - mean(current_data$cd4)) /
  (2*sd(current_data$cd4))
hist_data$age <- (hist_data$age - mean(hist_data$age)) /
  (2*sd(hist_data$age))
hist_data$cd4 <- (hist_data$cd4 - mean(hist_data$cd4)) /
  (2*sd(hist_data$cd4))
```

Then we define the parameters for the analysis.

```{r define_params, cache = TRUE}
# set parameters
delta0 <- 1
lambda0 <- 1
a0_seq <- seq(0, 1, length.out = 21)
data <- list(current_data, hist_data)
family <- binomial(link = "logit")
c0 <- 1
d0 <- 1
iter_warmup <- 1000
iter_sampling <- 2500
```

Lastly, we run the analysis and show the results.

```{r run_analysis, cache = TRUE}
post_samples <- samples_models(data = data,
                               outcome = "outcome",
                               family = family,
                               a0_seq = a0_seq,
                               c0 = c0,
                               d0 = d0,
                               delta0 = delta0,
                               lambda0 = lambda0,
                               iter_warmup = iter_warmup,
                               iter_sampling = iter_sampling,
                               num_cores = 14)
post_samples$df_post_ord
```